### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c76c179bbea5c8313459ae2046bec2ff4f52c611 🚀

Spillover risk is not unidimensional. From the standpoint of an animal
community, i.e.~a pool of possible hosts, there are a multiplicity of
ecological factors that come into play (Plowright et al. 2017). The
global richness of hosts is one such component commonly
mentioned/analysed (Anthony et al.~2017), but there is an argument to be
made that species who are not competent hosts of a specific virus genus
may not factor into this (Plowright et al. 2015), or that species who
are assumed to share viruses at different rates should be weighted
accordingly (Albery et al.~2020). In mammals, key functional traits (for
which phylogeny is a reasonable proxy) are determinants of the spillover
potential (Olival et al.~2017); these include, notably, TK. Finally,
especially when the pool of potential hosts spans the entire globe,
there may be local host pools that are highly unique; not having been
observed in other locations, these can act on the overall risk either by
providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid changes in specialism (Agosta et al. 2010). In the specific case
of generalist pathogens (as betacoronavirus clearly are), there is
conceptual and empirical support to the idea that these community- level
mechanisms are even more important in driving the overall risk (Power
and Mitchell 2004).

@@

@@

@@

>>>>

Spillover risk is not unidimensional. From the standpoint of an animal
community, \emph{i.e.} a pool of possible hosts, it is driven by a
multiplicity of factors (Plowright et al. 2017). The global richness of
hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
Anthony et al. 2017 for coronaviruses), but there is an argument to be
made that species who are not competent hosts of a specific virus genus
may not factor into this (Plowright et al. 2015), or that species who
are assumed to share viruses at different rates should be weighted
accordingly (Albery et al. 2020). In mammals, key functional traits (for
which phylogeny is a reasonable proxy) are determinants of the spillover
potential (Olival et al. 2017); these include, notably, body mass and
affinity for urban environments (Albery et al. 2022). Finally,
especially when the pool of potential hosts spans the entire globe,
there may be local host pools that are highly unique; not having been
observed in other locations, these can act on the overall risk either by
providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid changes in specialism (Agosta et al. 2010). In the specific case
of generalist pathogens (as betacoronavirus clearly are), there is
conceptual and empirical support to the idea that these community- level
mechanisms are even more important in driving the overall risk (Power
and Mitchell 2004).

@@

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
Albery, G. F. et al. 2020. Predicting the global mammalian viral sharing
network using phylogeography. - Nature Communications 11: 2260.

\leavevmode\hypertarget{ref-Albery2022UrbMam}{}%
Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
pathogens. in press.

\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
Anthony, S. J. et al. 2017. Global patterns in coronavirus diversity. -
Virus Evolution in press.

@@

\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
Olival, K. J. et al. 2017. Host and viral traits predict zoonotic
spillover from mammals. - Nature 546: 646--650.

@@

\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
Power, A. G. and Mitchell, C. E. 2004. Pathogen Spillover in Disease
Epidemics. - The American Naturalist 164: S79--S89.
